Is the Future of Authenticity In Our Heads?\\\medskip \large Moving Passthoughts From the Lab to the World

### 2017-04-07 paper.tex
revisions

% Created 2017-04-07 Fri 16:17
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:org87242b0}

@@

First, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
Through this work, we may build a better understanding of the statistical distribution of EEG signals that a person gives off during the course of their life. 
This statistical analysis will help us understand the space of possibile passthoughts. 
It will also help us estimate how easy or difficult passthoughts are to guess, or crack.

Second, we must understand the real usability properties of passthoughts.
While much work has claimed passthoughst to be more usable than traditional multi-factor authentication protocols,
as of yet, no work systematically evaluates this claim.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally.
Past work indicates that people believe EEG can reveal what someone is thinking or feeling, which could scare off wider adoption (Section 4).

The main goal of this paper is to motivate and contextualize immediate studies that can address the above issues.
Section 5 proposes two, complementary studies.
The first proposed study focuses on the use of passthoughts in a realistic, though lab-constrained setting.
By fixing some parameters (context of use, and feedback to the user), this study would allow a first look at how passthoughts might work
in a real-time application context.
The second proposed study focuses on people's longitudinal relationships with EEG in their everyday life.
This study focuses not on differences in behavior between experimental conditions, but on evolving attitudes (and changing signals) over time,
as a way of assessing both dynamic user attitudes, and the possible impact of shifting neural signals.

@@

\label{fig:earbud}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/passthoughts-diagram.png}
\caption{A passthought authenticator.}
\end{figure}

\section{Background}
\label{sec:org1286678}

@@

\subsection{One-step, multi-factor authentication}
\label{sec:orgc34f372}

@@

\subsection{Passthought authentication}
\label{sec:orga70a9dc}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:orge2da838}

@@

\cite{curranpassthoughts} was the first to merge in-ear EEG with passthought authentication,
 using a modified consumer grade EEG device with a single electrode, achieving approximately 80 percent authentication accuracy. 
Ongoing work from the same authors investigates the use of custom-fit earbuds with multiple embedded electrodes \ref{fig:earbud}.
Lending credibility to that study's claim that in-ear EEG could one day become feasible in consumer devices,
United Sciences recently announced a consumer "hearable'' (in-ear wearable) called The Aware, which will measure EEG from the ear, among other biometrics \cite{UnitedSciences}.

\subsection{Contending with mind-reading machines}
\label{sec:org463ed13}

@@

Crucially, there are some people who actually \emph{want} their minds measured, e.g. for self-reflection. Consider the Spire, a breath sensor that claims to divine, from a person's patterns of in-breaths and out-breaths, what the user is calm, focused, or tense \cite{SpireInc}. 
For the device to ``work,'' not only must these detected signals match with end-users' intuitions, but users must also believe that a device like the Spire has the power to measure and detect these phenomena, given breath as input \cite{Ali2014a}. 
In general, technologies that claim to ``measure the mind'' must rely on end-users to define the criteria by which systems are deemed effective, or accurate.

@@

\section{Diversity and crackability of passthoughts}
\label{sec:org8b8d857}

\section{User Perceptions of EEG}
\label{sec:orgd436f39}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:orgad4f714}

@@

\section{Two studies on passthoughts}
\label{sec:org0426e70}

This section proposes two studies on passthought authentication.

\subsection{A real-time passthought authenticator}
\label{sec:org85dc730}

@@

\subsubsection{The effect of feedback}
\label{sec:org63fcbe5}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:orgc3c28b6}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orgd647693}

@@

\subsubsection{A more diverse corpus}
\label{sec:org6aad8bd}

@@

\subsubsection{Usability and attitudes}
\label{sec:orgd8e2296}

By deploying a real sensing apparatus, be it a traditional consumer device such as the Muse \cite{Mihajlovic2015} 
or a more experimental piece of equipment such as an earbud \cite{UnitedSciences}, 
and having people record EEG data in their daily life, we could learn more about the interpretative qualities of these data  \cite{NafusDawn;Sherman2014}.
Such a study presents a dual opportunity to understand user beliefs in a rich, qualitative sense, while simultaneously collecting the large, diverse and longitudinal corpus of EEG signals necessary if we wish to stand a chance at decent classification accuracy in the wild.

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:orgd1ac1ca}

@@

\subsection{Privacy}
\label{sec:org3b2cea6}
One clear risk comes to user privacy, as it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could, at least theoretically, come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular datbase queries.

\subsection{Security}
\label{sec:org7d15f96}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\section{Further Future Directions}
\label{sec:org3b9b6f1}

@@

\subsection{Closed-loop (real-time) passthoughts}
\label{sec:orgc566ce1}
Future work on passthoughts should look at closed-loop, or online authentication systems,
in part to investigate the impact of human learning effects on passthought performance.
What effect does the feedback (of a successful or unsuccessful authentication attempt) have on the way that people perform their passthoughts?
Specific studies could, for example, provide false feedback in which passthought authentication appears to always either succeed or fail. In the always-fail condition, we might expect subjects to alter the way they perform the passthought across multiple attempts; data of how such a change occurs could enable us to pre-empt changes observable in the wild.

\subsection{Continuous authentication}
\label{sec:orge366961}

@@

\subsection{Organic passwords}
\label{sec:org4b5fad6}

If EEG signals are nonstationary (changing over time), passthoughts will require online machine learning to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as gait, which may change more slowly for individuals, if it changes at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Neuroscience of authentication}
\label{sec:orgeb5bc64}

@@

Such a scenario raises serious ethical, and legal questions. 
How does such a system conform to legal definitions of a person?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as I am aware, novel, and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this still-novel method for authentication.
\subsection{Mobile health}
\label{sec:org2f78f9b}
Neuroscience fuels some of the most chilling predictions in science fiction \cite{Welsh2011}.
It also stands for some of the greatest possible advances in medicine, mental health, and understanding of animal behavior (including our own).
By collecting unstructured or semi-structured EEG data in the wild, 
passthought systems could help build better BCIs \cite{Grierson2011a},
or as training data for larger systems of diagnosis or analysis. 
One ambitious goal is to detect or even predict seizures \cite{Mormann2006}.

@@

\subsection{Passthoughts by any other sensor?}
\label{sec:org5da47d1}

@@

\section{Conclusion}
\label{sec:org8594ecd}

In general, as sensors grow smaller and cheaper, devices more connected, and machine learning more sophisticated, 
people will build increasingly high-resolution models of human physiology ``in the wild.''
Passthoughts present just a microcosm of the good such advances might bring, 
along with some of the most pressing anxieties: 
What does pervasive physiological recording mean for our privacy, security, safety? 
The balancing act between these risks and opportunities will prove recurring theme for decades to come.
In the meantime, probing the outer limits of ubiquitous, pervasive sensing can shed light on both the good and bad that our near future may bring.

>>>>

% Created 2017-04-07 Fri 16:52
% Intended LaTeX compiler: pdflatex
\documentclass[sigconf]{acmart}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\input{authors}
\setcopyright{rightsretained}
\acmDOI{10.475/123_4}
\acmISBN{123-4567-24-567/08/06}
\acmConference[NSPW '17]{New Security Paradigms Workshop}{October 2017}{Islamorada, Florida, USA}
\acmYear{2017}
\copyrightyear{2017}
\acmPrice{15.00}
\date{\today}
\title{Is the Future of Authenticity In Our Heads?\\\medskip
\large Moving Passthoughts From the Lab to the World}
\hypersetup{
 pdfauthor={},
 pdftitle={Is the Future of Authenticity In Our Heads?},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 25.1.1 (Org mode 9.0.4)}, 
 pdflang={English}}
\begin{document}

@@

\section{Introduction}
\label{sec:orgad257ec}

@@

First, we must understand the real usability properties of passthoughts.
While much work has claimed passthoughst to be more usable than traditional multi-factor authentication protocols,
as of yet, no work systematically evaluates this claim.
The usability of passthoughts authentication will depend not only on its performance in ecologically valid contexts,
but also on people's attitudes about brainwaves, EEG, and biosensing generally (Section 3).

Second, we must test passthoughts in a variety of conditions: ambulatory settings, under different levels of stress, drowsiness, caffeine or alcohol, etc.
This statistical analysis will help us understand how passthoughts change over time.
It will also help us understand the space of possible passthoguhts (Section 4).

\label{fig:diagram}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./figures/passthoughts-diagram.png}
\caption{A passthought authenticator.}
\end{figure}

The main goal of this paper is to motivate and contextualize immediate studies that can address the above issues (Section 5).
The first study focuses on the use of passthoughts in a realistic, though lab-constrained setting, 
allowing a first look at how passthoughts might work in a real-time application context.
The second proposed study focuses on people's longitudinal relationships with EEG in their everyday life,
focusing not on behavioral differences experimental conditions, but on evolving attitudes (and signals) over longer timescales,
allowing assessment of both dynamic user attitudes, and the possible impact of shifting neural signals.

@@

\section{Background}
\label{sec:org4dc5ef5}

@@

\subsection{One-step, multi-factor authentication}
\label{sec:org416ad55}

@@

\subsection{Passthought authentication}
\label{sec:org9645597}

@@

\subsection{Passthoughts using in-ear EEG}
\label{sec:orgfd91d74}

@@

\cite{curranpassthoughts} was the first to merge in-ear EEG with passthought authentication,
 using a modified consumer grade EEG device with a single electrode, achieving approximately 80 percent authentication accuracy. 
Ongoing work from the same authors investigates the use of custom-fit earbuds with multiple embedded electrodes \ref{fig:earbud}.
Lending credibility to that study's claim that in-ear EEG could one day become feasible in consumer devices,
United Sciences recently announced a consumer "hearable'' (in-ear wearable) called The Aware, which will measure EEG from the ear, among other biometrics.

\subsection{Contending with mind-reading machines}
\label{sec:orgdc80023}

@@

Crucially, there are some people who actually \emph{want} their minds measured, e.g. for self-reflection. Consider the Spire, a breath sensor that claims to divine, from a person's patterns of in-breaths and out-breaths, what the user is calm, focused, or tense.
For the device to ``work,'' not only must these detected signals match with end-users' intuitions, but users must also believe that a device like the Spire has the power to measure and detect these phenomena, given breath as input \cite{Ali2014a}. 
In general, technologies that claim to ``measure the mind'' must rely on end-users to define the criteria by which systems are deemed effective, or accurate.

@@

\section{User Perceptions of EEG}
\label{sec:org30e9f5a}

@@

\subsection{What (do you think) EEG can reveal about a person?}
\label{sec:org4244390}

@@

\section{Diversity and crackability of passthoughts}
\label{sec:orgba47d6a}

\section{Two studies on passthoughts}
\label{sec:org31b107a}

This section proposes two studies on passthought authentication.
One study, a controlled, lab-based experiment, seeks to raise fundamental questions about how the feedback fo a real-time authentication system may affect the performance of passthoughts.
It also begins to address certain, limited questions around the shifting nature of neural signals.
The second study, a longitudinal deployment, seeks to collect a large and diverse corpus of EEG signals, while probing people's beliefs and attitudes about EEG and brainscanning in everyday life.
Together, the findings of these proposed studies could serve as a starting point for addressing the major questions raised in the prior sections.

\subsection{A real-time passthought authenticator}
\label{sec:orgf895ad6}

@@

\subsubsection{The effect of feedback}
\label{sec:orgc529a54}

@@

\subsubsection{Exploring continuous re-calibration}
\label{sec:org6714417}

@@

\subsection{A longitudinal study on brainwave monitoring}
\label{sec:orgc659ab1}

@@

\subsubsection{A more diverse corpus}
\label{sec:orgb90c729}

@@

\subsubsection{Usability and attitudes}
\label{sec:org4421f28}

By deploying a real sensing apparatus, be it a traditional consumer device such as the Muse \cite{Mihajlovic2015} 
or a more experimental piece of equipment such as an earbud,
and having people record EEG data in their daily life, we could learn more about the interpretative qualities of these data \cite{NafusDawn;Sherman2014}.
Such a study presents a dual opportunity to understand user beliefs in a rich, qualitative sense, while simultaneously collecting the large, diverse and longitudinal corpus of EEG signals necessary if we wish to stand a chance at decent classification accuracy in the wild.

@@

\section{Privacy, Security: Choices, Tradeoffs}
\label{sec:org05fd2a6}

@@

\subsection{Privacy}
\label{sec:org665d86b}
One clear risk comes to user privacy, as it is still not well understood what EEG signals might reveal about a person.
EEG signals that are not anonymized could, at least theoretically, come to be seen as private in the face of new methods of analysis.
(If your brainwaves can authenticate you, could they also uniquely identify you, even if your name is redacted?)
Differential privacy \cite{Dwork2014} presents one approach to dealing with the risk of privacy breaches with EEG signals.
By adding noise to datasets, differentially private databases can make strong guarantees about the likelihood of a de-anonymization attack on particular datbase queries.

\subsection{Security}
\label{sec:orgf04f8ff}
Device security presents another risk to passthought authentication.
Since EEG devices will transmit data, likely wirelessly \cite{Mihajlovic2015}, their data may be intercepted, depending on the security properties of the underlying transit protocol. 
When transferring authentication credentials in passthoughts, the ability to snoop on authentication attempts could present a dangerous attack vector.

@@

\section{Further Future Directions}
\label{sec:org8adb539}

@@

\subsection{Closed-loop (real-time) passthoughts}
\label{sec:orgedc89f4}
Future work on passthoughts should look at closed-loop, or online authentication systems,
in part to investigate the impact of human learning effects on passthought performance.
What effect does the feedback (of a successful or unsuccessful authentication attempt) have on the way that people perform their passthoughts?
Specific studies could, for example, provide false feedback in which passthought authentication appears to always either succeed or fail. In the always-fail condition, we might expect subjects to alter the way they perform the passthought across multiple attempts; data of how such a change occurs could enable us to pre-empt changes observable in the wild.

\subsection{Continuous authentication}
\label{sec:orgd885dff}

@@

\subsection{Organic passwords}
\label{sec:org6c324c8}

If EEG signals are nonstationary (changing over time), passthoughts will require online machine learning to maintain decent accuracy \cite{Vidaurre2006a}.
This feature of BCIs could have an unexpected benefit to security. 
If an individual's expression of their passthought in EEG is always changing, 
passthoughts themselves are effectively evergreen, automatically replaced or updated by nature of the authentication paradigm.
This feature could improve security, as an attacker able to compromise a passthought's EEG signature may not be able to log into the system in a few weeks time,
unless they are able to realistically mutate the signal over authentication attempts.
This feature of EEG also gives passthoughts a possible advantage over other methods for behavioral authentication, such as gait, which may change more slowly for individuals, if it changes at all.
Future work should investigate this claim, perhaps using a longitudinal corpus such as the one described above.
\subsection{Neuroscience of authentication}
\label{sec:orgf8773ab}

@@

Such a scenario raises serious ethical, and legal questions. 
How does such a system conform to legal definitions of a person?
Who is a person to make decisions for their future self?
What are possible vectors for abuse?
In any case, this property of an authentication is, as far as I am aware, novel, and should be considered as we learn more about the strengths, weaknesses, and particular affordances of this still-novel method for authentication.
\subsection{Mobile health}
\label{sec:orgb63fb70}
Neuroscience fuels some of the most chilling predictions in science fiction \cite{Welsh2011}.
It also stands for some of the greatest possible advances in medicine, mental health, and understanding of animal behavior (including our own).
By collecting unstructured or semi-structured EEG data in the wild, 
passthought systems could help build better BCIs \cite{Grierson2011a},
or as training data for larger systems of diagnosis or analysis. 
One ambitious goal is to detect or even predict seizures \cite{Mormann2006}.

@@

\subsection{Passthoughts by any other sensor?}
\label{sec:orgf9630d1}

@@

\section{Conclusion}
\label{sec:org3cc930b}

In general, as sensors grow smaller and cheaper, devices more connected, and machine learning more sophisticated, 
people will build increasingly high-resolution models of human physiology ``in the wild.''
Passthoughts present just a microcosm of the good such advances might bring, 
along with some of the most pressing anxieties: 
What does pervasive physiological recording mean for our privacy, security, safety? 
The balancing act between these risks and opportunities will prove recurring theme for decades to come.
In the meantime, probing the outer limits of ubiquitous, pervasive sensing can shed light on both the good and bad that our near future may bring.
\cite{Gaver1999}
